Pre-trained encoders for dialogue: The case of laughter

### 2019-12-09 reports/ACL2020 Paper/acl2020-paper.tex
create background section
Some more minor fixes

\begin{document}
\maketitle
\begin{abstract}
  The degree to which language models pre-trained on text data are useful for dialogue is an open question.
  In this paper, we investigate the usefulness of BERT for dialogue act recognition.
  We are interested in both how useful the pre-training procedure is and 
  how well the model adapts to the dialogue domain.
  To confirm that the fine-tuned model learns to represent dialogical features, 
  we look at how it uses laughter, a phenomenon specific to dialogue.
\end{abstract}

@@

In this work, we investigate whether BERT \citep{devlinBERTPretrainingDeep2018}, a well-know pre-trained language model, is useful for dialogue.
We use dialogue act recognition  as a proxy task, since both the internal content of contributions, and their sequential structure have bearing on the task.
In this work we also show that laughs can be helpful to disambiguate a dialogue act.

@@

@@

@@

Confusion matrices (figure \ref{fig:swda-cm}) provide some food for thought.
% TODO: explain why most of misclassifications are sd (majority class)
One of the most obvious things that can be observed is rhetorical questions: supposedly, laughter can be used as a device to cancel seriousness \citep{ginzburg2015understanding} of the question.
Therefore it can help to disambiguate such questions like the one we show in example in table~\ref{table:example-qh}.
\begin{table}
      \small
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
        \multicolumn{3}{c}{\emph{(Talking about hobbies)}}\\
    B   & sd	&  Um, as far as spare time, they talked about, \\
    B	& \% & I don't, + I think, \\
    B	& qh & who has any spare time \texttt{<laughter>}? \\
    A	& x & \texttt{<laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw3735). B's contribution \emph{qh} (Rhetorical question) is misinterpreted as \emph{qw} (Wh-question) by the BERT model without laughs in training data. }
  \label{table:example-qh}
\end{table}

  \begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{img/swda-cm.pdf}
  \caption{Confusion matrices for \texttt{BERT-NL} (top) vs \texttt{BERT-L} (bottom); SWDA corpus. Solid line shows the improvement in classification of rhetorical questions.}
    \label{fig:swda-cm}
  \end{figure}

@@

%\subsection{Experiment 4: Cross-domain pre-training (if time permits)} \label{sec:experiment4} % Bill
%  - cross-domain pre-training
\section{Related work} % add references first
For the background:

Dialogue act recoginition
\begin{itemize}
  \item \citet{botheContextbasedApproachDialogue2018} -- uses an RNN for context (recently SOTA on SWDA)
  \item \citet{pragstVectorRepresentationUtterances2018} -- dialogue vector models. Uses vector-manual DA tag correspondance to argue for semantic relevance of their vector models (related to how we use DA tagging to verify BERT usefulness)
  \item \citet{cerisaraEffectsUsingWord2vec2017} -- standard pre-trained word embeddings do not help for DAR in any of three tested languages (English, French, Czech)
  \item \citet{kalchbrennerRecurrentConvolutionalNeural2013} -- DAR model similar to our architecture. Uses an RNN ``discourse model'' and a CNN similar to our baseline encoder
  \item \citet{khanpourDialogueActClassification2016} -- Deep LSTM utt representation. No notion of context. Finds word vectors are useful.
  \item \citet{tranPreservingDistributionalInformation2017} -- Sequential DA prediction with uncertainty propegation
  \item \citet{chenDialogueActRecognition2017} -- SOTA? SWDA DAR using attention and CRF over utterances
  \item \citet{ortegaNeuralbasedContextRepresentation2017} -- SWDA DAR with attention-based context representations and CNN utterance representations
  \item \citet{shenNeuralAttentionModels2016} -- Uses attention for utt encoding in DAR. No notion of context, I think.
  \item \citet{tranHierarchicalNeuralModel2017} -- RNNs at both the utteranec and dialogue level. Also attention. Very similar to our architecture.
\end{itemize}

Pre-training/transfer learning for dialogue
\begin{itemize}
  \item \citet{mehriStructuredFusionNetworks2019} -- ``Structured fusion networks''. Traditional dialogue system design with neural models.
  \item \citet{chenSemanticallyConditionedDialog2019a} -- uses BERT for response generation
  \item \citet{baoPLATOPretrainedDialogue2019} -- pre-training for dialogue generation
  \item \citet{mehriPretrainingMethodsDialog2019} -- looks at various dialogue tasks using pre-training, including DAR and using BERT
  \item \citet{vigComparisonTransferLearningApproaches} -- Transfer learning for response selection
\end{itemize}

>>>>

\begin{document}
\maketitle
\begin{abstract}
  In this paper, we investigate the usefulness of BERT for dialogue act recognition.
  We are interested in both how useful the pre-training procedure is and 
  how well the model adapts to the dialogue domain.
  To confirm that the fine-tuned model learns to represent dialogical features, 
  we look at how it uses laughter, a phenomenon specific to dialogue.
\end{abstract}

@@

We investigate whether BERT \citep{devlinBERTPretrainingDeep2018}, a well-know pre-trained language model, is useful for dialogue.
We use dialogue act recognition  as a proxy task, since both the internal content of contributions, and their sequential structure have bearing on the task.
We also show that laughs can be helpful to disambiguate a dialogue act.

\section{Background}

@@

\begin{itemize}
  % related because of the DAR sequence model
  \item \citet{kalchbrennerRecurrentConvolutionalNeural2013} -- DAR model similar to our architecture. Uses an RNN ``discourse model'' and a CNN similar to our baseline encoder
  \item \citet{botheContextbasedApproachDialogue2018} -- uses an RNN for context (recently SOTA on SWDA)
  \item \citet{tranHierarchicalNeuralModel2017} -- RNNs at both the utteranec and dialogue level. Also attention. Very similar to our architecture.
  \item \citet{tranPreservingDistributionalInformation2017} -- Sequential DA prediction with uncertainty propegation
  % cf.
  \item \citet{ortegaNeuralbasedContextRepresentation2017} -- SWDA DAR with attention-based context representations and CNN utterance representations
  % related because of the utt encoder model
  \item \citet{khanpourDialogueActClassification2016} -- Deep LSTM utt representation. No notion of context. Finds word vectors are useful.
  \item \citet{shenNeuralAttentionModels2016} -- Uses attention for utt encoding in DAR. No notion of context, I think.
  % related because of methodology
  \item \citet{pragstVectorRepresentationUtterances2018} -- dialogue vector models. Uses vector-manual DA tag correspondance to argue for semantic relevance of their vector models (related to how we use DA tagging to verify BERT usefulness)
  % ???
  \item \citet{chenDialogueActRecognition2017} -- SOTA? SWDA DAR using attention and CRF over utterances
\end{itemize}

@@

\begin{itemize}
  \item \citet{cerisaraEffectsUsingWord2vec2017} -- standard pre-trained word embeddings do not help for DAR in any of three tested languages (English, French, Czech)
    % BERT for dialogue (not nesc. DAR)
  \item \citet{chenSemanticallyConditionedDialog2019a} -- uses BERT for response generation
  \item \citet{mehriPretrainingMethodsDialog2019} -- looks at various dialogue tasks using pre-training, including DAR and using BERT
  \item \citet{baoPLATOPretrainedDialogue2019} -- pre-training for dialogue generation
  \item \citet{vigComparisonTransferLearningApproaches} -- Transfer learning for response selection
\end{itemize}

\paragraph{Contributions}
The main results of this paper have to do with to the usefulness of language model fine-tuning for DAR, the impact of laughter on DAR, and the relationship between the two.
\begin{itemize}
  \item Laughter is useful for dialogue act recognition, and its impact varies across different dialogue acts (\S\ref{sec:experiment1}).
  \item During fine-tuning, BERT learns to represent laughter, a dialogical feature not seen in pre-training (\S\ref{sec:experiment2} and \S\ref{sec:experiment3}).
  \item Standard BERT pre-training is useful for DAR, but the model performs poorly without fine-tuning (\S\ref{sec:experiment2})
  \item Further pre-training with in-domain data shows promise for dialogue (\S\ref{sec:experiment3}), but further investigation with larger dialogue corpora is required.
\end{itemize}

@@

Confusion matrices (figure \ref{fig:swda-cm}) provide some food for thought. Most of the misclassifications fall into the majority classes, such as \emph{sd} (Statement-non-opinion), in the left edge of the matrix. However there are some important exceptions, such as \emph{rhetorical questions}, that are misclassified as other forms of questions due to their surface question-like form.
Importantly, laughter helpes to classify rhetorical questions correctly, this is because in a conversation it can be used as a device to cancel seriousness \citep{ginzburg2015understanding}, in this case, of a question.
Therefore questions, like the one we show in the example in table~\ref{table:example-qh}, are easier to disambiguate with the help of laughter.
\begin{table}
      \small
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
        \multicolumn{3}{c}{\emph{(Talking about hobbies)}}\\
    B   & sd	&  Um, as far as spare time, they talked about, \\
    B	& \% & I don't, + I think, \\
    B	& qh & who has any spare time \texttt{<laughter>}? \\
    A	& x & \texttt{<laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw3735). B's contribution \emph{qh} (Rhetorical question) is misinterpreted as \emph{qw} (Wh-question) by the BERT model without laughs in training data. }
  \label{table:example-qh}
\end{table}

  \begin{figure}
  \centering
  \includegraphics[width=1.0\linewidth]{img/swda-cm.pdf}
  \caption{Confusion matrices for \texttt{BERT-NL} (top) vs \texttt{BERT-L} (bottom); SWDA corpus. Solid line shows the improvement in classification of rhetorical questions.}
    \label{fig:swda-cm}
  \end{figure}

@@

%\subsection{Experiment 4: Cross-domain pre-training (if time permits)} \label{sec:experiment4} % Bill
%  - cross-domain pre-training
